Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-01-12 main.tex
Updates from Overleaf

The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 for reported cases and deaths from COVID-19. See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data. We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}). We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values. In addition, we filtered out erroneous forecasts defined by any of the conditions listed in Table \ref{tab:erroneous}. Those forecasts were removed in order to be better able to illustrate the effects of the log-transformation on scores and eliminating distortions caused by outlier forecasters. All predictive quantiles were truncated at 0. We applied the log-transformation after adding 1 to all predictions and observed values. Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

@@

Show and discuss Fig 7 here? CHECK AGAIN

\subsection{Descriptive analysis of observed values and scores obtained on the natural and logarithmic scales CHECK AGAIN}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-transformation-regression.png}
    \caption{Relationship between median forecasts and scores. Black dots represent WIS values for two-week ahead predictions of the EuroCOVIDhub-ensemble. Shown in red are the regression lines discussed in \ref{sec:regression} and \ref{tab:HUB-regression}. A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values. B: Same as A, with scores obtained after applying a log-transformation to the data. C: Same as A, with scores obtained after applying a square-root-transformation to the data.} 
    \label{fig:HUB-regression}
\end{figure}

\subsection{Regression analysis to determine the variance-stabilizing transformation}

@@

To check this, we ran the regressions
\begin{equation*}
    \text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
\end{equation*}
where $F_{\log}$ is the predictive distribution for $\log(y)$, and
\begin{equation*}
    \text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
\end{equation*} 
where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale. As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate small targets lead to larger average CRPS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$). The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}. It appears that the log transformation is most suitable to stabilise the WIS values in this empirical example.

@@

For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A). When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably. The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}B) and possibly tends to increase with increasing forecast horizon.

When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge. Missing the peak, i.e. predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B). On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D). We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}B is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases. 
Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low. On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)). One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F). On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.

@@

>>>>

The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 5th of December 2022 for reported cases and deaths from COVID-19. See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data. We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}). We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values. In addition, we filtered out erroneous forecasts defined by any of the conditions listed in Table \ref{tab:erroneous}. Those forecasts were removed in order to be better able to illustrate the effects of the log-transformation on scores and eliminating distortions caused by outlier forecasters. All predictive quantiles were truncated at 0. We applied the log-transformation after adding 1 to all predictions and observed values. Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

@@

When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge. Missing the peak, i.e. predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B). On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D). Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low. On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)). One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F). On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.

@@

\subsection{Regression analysis to determine the variance-stabilizing transformation}
\label{sec:HUB-regression}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-transformation-regression.png}
    \caption{Relationship between median forecasts and scores. Black dots represent WIS values for two-week ahead predictions of the EuroCOVIDhub-ensemble. Shown in red are the regression lines discussed in Section \ref{sec:HUB-regression} shown in Table \ref{tab:HUB-regression}. A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values. B: Same as A, with scores obtained after applying a log-transformation to the data. C: Same as A, with scores obtained after applying a square-root-transformation to the data.} 
    \label{fig:HUB-regression}
\end{figure}

To check this, we ran the regressions
\begin{equation*}
    \text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
\end{equation*}
where $F_{\log}$ is the predictive distribution for $\log(y)$, and
\begin{equation*}
    \text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
\end{equation*} 
where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale. As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate small targets lead to larger average CRPS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$). The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}. In this empirical example, the log transformation helps (albeit not perfectly), to stabilise WIS values and achieves this better suited than the square-root transformation.

@@

For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A). When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably. The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}B) and increases noticeably with increasing forecast horizon.

@@

We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}B is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases.
